---
title: "flight_delay"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{flight_delay}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(linereg)
library(caret)
library(rminer)
```
**Description** 

In this vignette we will try to predict the delay of each flights using our own ridgereg() function from linreg package. The data comes from nycflights13 package.

**Application** 

1. Geting weather dataset and the flights dataset from the nycflights13 package and removing the eventual variables which do not have a predictive value.
```{r,include=TRUE,eval=FALSE,echo=TRUE,results='hide'}
#getting flight dataset from nycflights13 library and removing rows 
#in which na is present in dep_delay and arr_delay columns
df_flights<- nycflights13::flights
df_flights <- df_flights[!is.na(df_flights$dep_delay), ]
df_flights <- df_flights[!is.na(df_flights$arr_delay), ]

#getting the arr_delay which is greater than 0
library(dplyr)
df_flights<-  filter(df_flights ,arr_delay>0)

#creating a new dataframe carrier_group grouping by origin, tailnum, day
#hour and summarising by the arr_delay,dep_delay,distance and air_time columns
carrier_group <- df_flights %>%
    group_by( origin,tailnum, month, day, hour) %>%
  summarise(arr_delay = arr_delay,
            dep_delay=dep_delay,
            distance=distance,
            air_time=air_time)

#removing rows based on na values in the below columns
carrier_group<-carrier_group[!is.na(carrier_group$dep_delay), ]
carrier_group<-carrier_group[!is.na(carrier_group$arr_delay), ]
carrier_group<-carrier_group[!is.na(carrier_group$tailnum), ]
carrier_group<-carrier_group[!is.na(carrier_group$month), ]
carrier_group<-carrier_group[!is.na(carrier_group$day), ]
carrier_group<-carrier_group[!is.na(carrier_group$hour), ]
carrier_group<-carrier_group[!is.na(carrier_group$distance), ]
carrier_group<-carrier_group[!is.na(carrier_group$air_time), ]
```
2. Add extra weather data from the weather dataset and organize it to get the final data frame.
```{r,include=TRUE,eval=FALSE,echo=TRUE,results='hide'}
#creating new dataframe weat by taking the required columns from weather dataset

weat <-  nycflights13::weather %>%
    group_by( origin,month, day, hour) %>%
  summarise(temp = mean(temp),
            dewp=dewp,
            humid=humid,
            wind_dir=wind_dir,
            speed=wind_speed,
            visib=visib,
            wind_gust=wind_gust,
            precip=precip)

#removing rows based on na values in the below columns
weat <- weat[!is.na(weat$wind_gust), ]
weat<-weat[!is.na(weat$month), ]
weat<-weat[!is.na(weat$day), ]
weat<-weat[!is.na(weat$hour), ]
weat<-weat[!is.na(weat$precip), ]
weat<-weat[!is.na(weat$visib), ]
weat<-weat[!is.na(weat$speed), ]
weat<-weat[!is.na(weat$wind_dir), ]
weat<-weat[!is.na(weat$humid), ]
weat<-weat[!is.na(weat$dewp), ]
weat<-weat[!is.na(weat$temp), ]


#left joining weat dataframe to get our final carrier_group dataframe
carrier_group <- weat %>%
  left_join(carrier_group, weat, by = c('month'='month', 'day'='day', 'hour'='hour','origin'='origin'))

#removing rows based on na values in the below columns
carrier_group<-carrier_group[!is.na(carrier_group$dep_delay), ]
carrier_group<-carrier_group[!is.na(carrier_group$arr_delay), ]
carrier_group<-carrier_group[!is.na(carrier_group$tailnum), ]
carrier_group<-carrier_group[!is.na(carrier_group$month), ]
carrier_group<-carrier_group[!is.na(carrier_group$day), ]
carrier_group<-carrier_group[!is.na(carrier_group$hour), ]
carrier_group<-carrier_group[!is.na(carrier_group$distance), ]
carrier_group<-carrier_group[!is.na(carrier_group$precip), ]
carrier_group<-carrier_group[!is.na(carrier_group$visib), ]
carrier_group<-carrier_group[!is.na(carrier_group$speed), ]
carrier_group<-carrier_group[!is.na(carrier_group$wind_dir), ]
carrier_group<-carrier_group[!is.na(carrier_group$humid), ]
carrier_group<-carrier_group[!is.na(carrier_group$dewp), ]
carrier_group<-carrier_group[!is.na(carrier_group$temp), ]
carrier_group <- carrier_group[!is.na(carrier_group$wind_gust), ]
carrier_group<-carrier_group[!is.na(carrier_group$air_time), ]


#factorizing origin and tailnum categorical data
carrier_group$origin = factor (carrier_group$origin,
                               levels =c('EWR','JFK','LGA'),
                               labels=c(1,2,3))
carrier_group$tailnum<-as.factor(carrier_group$tailnum)
carrier_group[,c('tailnum')]<-sapply(carrier_group[,c('tailnum')],unclass)

```
3. Using the caret package to divide the fight dataset into three sets: test, train and validation.
```{r,include=TRUE,eval=FALSE,echo=TRUE,results='hide'}
#splitting the data into train, test and validation datasets
train_idx<-  createDataPartition(carrier_group$arr_delay,p=0.95,list=F)
test<-carrier_group[-train_idx,]

val_idx<-  createDataPartition(carrier_group$arr_delay[train_idx],p=0.16,list=F)
valid<-carrier_group[val_idx,]
train<-carrier_group[-val_idx,]
```
4. Training ridge regressions models for different values of λ and evaluate the root mean squared error on the validation set. And then to find an optimal value for λ.
```{r,include=TRUE,eval=FALSE,echo=TRUE}
rmse_vec<-c() # root mean squared errors 
r2_vec<-c() # R squared
mse_vec<-c() # mean squared errors
lamda<-seq(0.0001,0.1,by=0.0001)

#getting lambda value by predicting on validation dataset after tarining on dataset
for (i in 1:length(lamda)){
  
  model_lamda<-ridgereg(formula=arr_delay~.,data=train,lamda=i)
  ridge_pred_l<-model_lamda$predict(valid)
  rmse_ridge<- mmetric(valid$arr_delay,ridge_pred_l,"RMSE")
  mape_ridge<- mmetric(valid$arr_delay,ridge_pred_l,"MAPE")
  r2_ridge<- mmetric(valid$arr_delay,ridge_pred_l,"R2")
  rmse_vec<-c(rmse_vec,rmse_ridge)
  r2_vec<-c(r2_vec,r2_ridge)
  mape_vec<-c(mse_vec,mape_ridge)
}
idx<-which(rmse_vec==min(rmse_vec))
plot(lamda,rmse_vec)
```


5. using λ value to predict the test set and reporting the RMSE of the predicted model.
```{r,include=TRUE,eval=FALSE,echo=TRUE,results='hide'}
# we can see that rmse is very minimum that is around 0.0001

#Predicting the values of test data by training the model
model<-linereg::ridgereg(formula=arr_delay~.,data=train,lamda=0.0001)
model
pred<-model$predict(test)
```

RMSE value on predicting test data is shown below, prediction and RMSE value depends on selecting features and amount of data, but we can see the rmse values of test data is slightly higher than training data but lower than validation data this confirms our model is not overfitted

```{r,include=TRUE,eval=FALSE,echo=TRUE}
rmse<- mmetric(test$arr_delay,pred,"RMSE")
rmse

```