---
title: "linreg"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{linreg}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
#library(linereg)
```

**Description**  


The linreg package uses ordinary least squares method to calculate best fitted values, estimators(coefficients), residuals,variance of the regression coefficients residual variance, t values and P values.
  
  
**Installation** 


The Package can be downloaded from.
    
    
    devtools::install_github("yemimorrison/linereg")
    
    
**Vignette**  


After installation run the code below to find vignettes.
    
    
    browseVignettes('linereg')
    
    
**Attributes defined in the Class **

    Linreg Class contains following attributes:
        *formula (formula). Dependent and independent columns in data frame
        *data (data.frame). Data frame for linear models
        *coeffs (Vector). Estimators matrix converted to vector
        *pred_val (matrix). Estimated dependent values
        *e (matrix). Error calculated from actual minus calculated dependent varaiable
        *deg_f (numeric). degrees of freedom
        *var_e (matrix). Residual variance
        *var_coeffs (vector). variance of the regression coefficients
        *t_value (matrix). T value
        *df_name (character). Data Frame name


**Methods**


1.print()  

  Prints out the coefficients.


2.plot()  

  Plots the following **Residuals vs Fitted values** and **Standardized error vs Fitted values** plots using ggplot2


3.resid()  

  This method returns the vector of residuals(errors).


4.coef()  

  Returns the coefficients as a named vector.


5.summary()  

  Summary returns a matrix of Estimate, Standard Error, t value, P value, Residual standard error and degrees of freedom.




**Examples**   

    
    1) linreg_obj<-linreg$new(formula=Petal.Length ~ Species,data=iris)
    linreg_obj$print()
  
    call:
    linreg(formula = Petal.Length ~ Species, data = iris)

    Coffiecients:
      (Intercept)  Speciesversicolor   Speciesvirginica  
            1.462              2.798              4.090 
            
    
    
    2) linreg_obj<-linreg$new(formula=Petal.Length ~ Species,data=iris)
    linreg_obj$summary()
    
    call:
    linreg(formula = Petal.Length ~ Species, data = iris)

    Coffiecients:
                  Estimate Std.Error t value Pr(>|t|)                
    (Intercept)       1.462    0.06086   24.02   9.3030515228435e-53  ***
    Speciesversicolor 2.798    0.08607   32.51   5.25458742021361e-69 ***
    Speciesvirginica  4.09     0.08607   47.52   4.106138619051e-91   ***

    Residual standard error: 0.4318 on 147 degrees of freedom
    



